Predict Class of Exercise
========================================================
## Abstract
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this analysis, we choose 18 kind of features to predict which exercise class they do, 
and train using random forest (RF) method.
We find that the expected error rate is about 0.3% in test sample.
Therefore, we can predict which class of exercise they do.

## Prerequisite

```{r}
#code is commented because we downloaded the data once
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="pml-training.csv", method="curl");
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="pml-testing.csv", method="curl");
```
```{r}
library(caret)
```

## Data Preprocessing and Analysis

Loading data and creating partition of training set into two
- to train model
- to cross validate

Let's summerize training set picking correct and suitable features to predict

```{r,echo=TRUE}
raw_training <- read.csv(".//data//pml-training.csv");
testing  <- read.csv(".//data//pml-testing.csv");

set.seed(1)
inTrain <- createDataPartition(y=raw_training$classe,p=0.6, list=F)
training <- raw_training[inTrain, ]
cv       <- raw_training[-inTrain, ]

summary(training)

```

few patterns of training set that we have for training model.

```{r}
p1<-qplot(new_window, num_window, data=training)+facet_grid(.~classe)
p1
p2<-qplot(yaw_arm, total_accel_arm,data=training)+facet_grid(.~classe)
p2
p3<-qplot(roll_arm, pitch_arm,data=training)+facet_grid(.~classe)
p3
p4<-qplot(roll_belt, pitch_belt, data=training)+facet_grid(.~classe)
p4
```

Features that we select for prediction

```{r}
input_vars_list <- c("new_window", "num_window", "roll_belt", "pitch_belt", "yaw_belt", 
  "total_accel_belt", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", 
    "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell",
    "roll_forearm", "pitch_forearm", "yaw_forearm", 
    "total_accel_forearm");
	
exp_input <- function(x){
  res = x[1];
    for (i in 2:length(x)){
            res <- paste(res, " + ", x[i], sep="");
    }
    return(res);
}

missClass <- function(values, prediction){
  sum( prediction != values )/length(values)
}

input_vars <- exp_input(input_vars_list)
	
```	

Creating model by random forest.
```{r}

set.seed(2);

st <- Sys.time();
modFit <- train( eval(parse(text=paste("classe ~", input_vars, sep="")))
  ,data=training, method="rf")
et <- Sys.time();
et-st
```

##Results
check out error rate for training set and cross validation set

```{r}
missClass(training$classe, predict(modFit, training))

missClass(cv$classe, predict(modFit, cv))

```

Lets predict test samples using our model.

```{r}
# test sample
answers <- predict(modFit, testing)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
